

---
title: "Lab 3 Report"
author: "Jonathan Kanyinda"
date: "`r format(Sys.Date())`"
output: github_document
---

Instructions for this lab report can be found on [Your assignment section of Lab Assignment 3 introduction to ggplot2](https://biol275-msum.github.io/introduction-to-ggplot2.html#your-assignment) on the lab web site. 



## Fireflies

> A. Insert an R code chunk and create a graph depicting the frequency distribution of the 35 mass measurements. It should have legible text and appropriate axis labels.

```{r}
library(tidyverse)

library(auk)                          # load the auk package
birds <- ebird_taxonomy %>%           # start with the ebird_taxonomy data
  as_tibble() %>%                     # tibbles print better in the console
  filter(category == "species")       # remove non-species taxa


fireflies_data <- read_csv("https://whitlockschluter.zoology.ubc.ca/wp-content/data/chapter02/chap02q19FireflySpermatophoreMass.csv")

ggplot(data = fireflies_data) +
  geom_histogram(mapping = aes(x = spermatophoreMass), binwidth = 0.01,
                 boundary = 0, closed = "left", 
                 fill = "#C5351B", color = "black") +
  labs(x = "Spermatophore Mass (mg)", y = "Frequency (number of fireflies)") +
  scale_y_continuous(expand = expansion(mult = 0))+
  theme_classic() +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black", size = rel(1))
  )
  
```



\_\_\_

> B. What kind of graph did you create in part (A) and why?

The graph is a histogram. This is because histograms are used to show the distribution 
of a single numerical variable.

\_\_\_

> C. Describe the shape of the frequency distribution. What are its main features?

The shape is bimoda. However there is an overall bell shaped form for the data between 0 and 10.

\_\_\_

> D. What term would be used to describe the largest measurement in the frequency distribution? You can refer to the lecture slides, your notes, or the textbook.

The mode
\_\_\_

## Bird orders

I read the data using the code provided by the instructor. The dataset is from the auk package and was originally named ebird_taxonomy. I converted it to a tibble, removed non-species taxa, and named the new dataset birds.

```{r}
library(auk)                          # load the auk package
birds <- ebird_taxonomy %>%           # start with the ebird_taxonomy data
  as_tibble() %>%                     # tibbles print better in the console
  filter(category == "species")       # remove non-species taxa
```

> E. How many bird species are in the new birds dataset? How did you find out?

There are 10721 specis of birds. This is found by creating the table of the different scientific names and the tables displayed 10721 rows, each one representing a species

\_\_\_

> H. How many orders are there? You can get this right by counting on the graph, but you will only get full points if you use some code to figure it out.

There were 41 orders and this was found by creating the contangency table which displayed 10 rows and a message saying that there was 31 more rows which gave a total of 41 rows each of which represented an order
\_\_\_

> F. Insert an R code chunk and create a graph depicting the distribution of orders in the birds dataset. Sort the orders with the most frequent on the left. It should have legible text and appropriate axis labels.

```{r}

# load auk package

library(auk)                          # load the auk package
birds <- ebird_taxonomy %>%           # start with the ebird_taxonomy data
  as_tibble() %>%                     # tibbles print better in the console
  filter(category == "species")       # remove non-species taxa

# Final graph 

ggplot(data = birds) +
  geom_bar(mapping = aes(x = fct_infreq(order)), fill = "#C5351B", 
           width = .8) +
  labs(x = "Order", y = "Frequency (number of species)") +
  #scale_y_continuous(limits =  c(0, NA), expand = expansion(mult = 0)) +
  theme_classic(base_size = 12) +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black", size = rel(.6)),
    axis.text.x = element_text(angle = 70 , hjust = 1),
    axis.ticks.x = element_blank()
  ) + 
  scale_y_log10(expand = expansion(mult = 0))

```




\_\_\_

> G. What kind of graph did you create? Why?

I created a bar graph. This was because the distribution of a categorical variable is best visualized using a bar graph.

\_\_\_

## Links to peer review

I reviewed Nathan Snell's lab report at this URL: https://github.com/BIOL275-MSUM/lab-3-intro-to-ggplot-NathanSnell1/issues/1

I reviewed Tyler Edvall's lab report at this URL: https://github.com/BIOL275-MSUM/lab-3-intro-to-ggplot-tdedvall/issues/3

Haider Ali reviewed my lab report at this URL: https://github.com/BIOL275-MSUM/lab-3-intro-to-ggplot-Jonathan-kanyinda/issues/1

## Session Info

Here is my session information:

```{r}
sessioninfo::session_info()
```
